Continuous voting by approval and participation indexes

### 2015-04-24 ApresentacoesArtigos/collectiveRanking/paper.tex
repassada ateh discussao

\title{Approval and participation indexes for a voting process} % For titles, only capitalize the first letter

@@

\begin{abstract}
In finding the adequate way to prioritize proposals, the Brazilian participation community agreed about an approval index and a participation index.
Both practice and literature is constantly handled by the experts involved, and the formalization of such model and indexes seems novel.
Also, the relevance of this report is strengthened by the nearby use of these indexes, by the Brazilian General Secretariat of the Republic, to raise and prioritize proposals about public health care in open processes and assist federal public managers.
\end{abstract}

@@

\section{}
%------------------------------------------------
% aqui já mando toda intro com as infos que cavei ontem.
Online decision making is a kind of recommendation system with special appeal for online social participation and electronic governments. This poses challenges on the design of such
processes regarding validity, security and the adequate indicators. Indeed,
the processes themselves vary, and the fact that the ranking indexes presented here seem not to be
formalized and published is an evidence that such online decision is very recent phenomena.

The main contribution of this report is the indexes for
an online voting process~\cite{issue1,issue2,tabRonald} with the following characteristics:
\begin{itemize}
    \item proposals might be inserted by voters after the voting phase started.
    \item Voting might be extended as a permanent social participation instance. In other words, voting on and adding new proposals might be open continuously.
    \item A proposal is presented to a voter one by one as random outcomes of all proposals.
    \item Each vote might be of one and only type among: ``approve'', ``disapprove'' and ``indifferent''.
    \item Voters vote without authentication.
\end{itemize}

This setting requires care about security and validity.
Some of which are:
\begin{itemize}
	\item adequate estimates of threshold for statistical validity of the ranking.
    \item Keeping the IP address of voters to ease detection of automated and other fraudulent efforts.
    \item The use of the outcomes from the voting process. This requires probing the survey being conducted and its purposes. The indexes here presented targets the deliverance of indicatives for the Brazilian federal government about the most important health care proposals. Given the unauthenticated voting, the outcomes might be regarded as reference rankings if data is minimally shared and checked for inadequate data entry (such as voting by automated scripts or one persistent participant).
\end{itemize}

@@

The approval index $\alpha_i$ and the participation index $\gamma_i$ of the proposal $i$ was defined as:
\begin{equation}
\begin{split}
\alpha_i=\frac{v^+_i-v^-_i}{\eta_i} \\
\gamma_i=\frac{v^+_i+v^-_i}{\eta_i}
\end{split}
\end{equation}
where $v^+_i$, $v^-_i$ and $\eta_i$ are approval count, disapproval count and exhibition count, respectively.
Note that $\alpha_i\in[-1,1]$, $\gamma_i\in[0,1]$, and $v^o_i=\eta_i-v^+_i-v^-_i$ is the count of the ``indifferent'' manifestations received by proposal $i$.
Also, such $\alpha_i$ and $\gamma_i$ indexes are expected, for each proposal $i$, to be constants  plus a sampling estimate error that should be smaller as $\eta_i$ raises. This error is thought to be acceptable if $\eta_i$ is above a threshold $\overline{\eta}$ established by the participation community and public managers.
As an initial decision, the staff agreed to use $\overline{\eta}$ as to select $10-20\%$ of all proposals.
A threshold $\overline{\gamma}$ can be used as a required level of engagement for proposals to be relevant,
while the threshold $\overline{\alpha}$ is used to classify the outcome
as ``approved'',``disapproved'' and ``clash''. More specifically:

@@

If a proposal is both sampled and approved, than it is prioritized.
The coherent values of $\overline{\alpha_i}=0.5$ (or $1/3$)
and $\overline{\gamma_i}=0.5$ were chosen as standards of the decision model. 
These are likely to change with implementation and management.

@@

@@

A Brazilian solution, used in diverse software and specially important as the output of a nation-wide decision making need, is the Agora Algorithm~\cite{agora}. It presents a decision procedure in phases (agenda proposition, deliberations proposition and commenting, voting) with resolution outcomes.
Although coherent, this framework requires authentication and might need experimentation and tuning so to achieve the desired use with more than dozens or a few hundreds of participants.

>>>>

\title{Continuous voting by approval and participation indexes} % For titles, only capitalize the first letter

@@

\begin{abstract}
In finding the adequate way to prioritize proposals, the Brazilian participation community agreed about an approval index and a participation index.
Both practice and literature is constantly handled by the experts involved, and the formalization of such model and indexes seems novel.
Also, the relevance of this report is strengthened by the nearby use of these indexes by the Brazilian General Secretariat of the Republic to raise and prioritize proposals about public health care in open processes.
\end{abstract}

@@

\section{}
%------------------------------------------------
% aqui já mando toda intro com as infos que cavei ontem.
Online decision making is a kind of recommendation system with special appeal for online social participation and electronic governments. This poses challenges on the design of such
processes regarding validity, security and the adequate indicators. Indeed,
the processes themselves vary, and the fact that the ranking indexes presented here seem not to be
formalized and published is an evidence that such online decision making is very recent phenomena.

The main contribution of this report is the indexes for
an online voting process~\cite{issue1,issue2,tabRonald} with the following characteristics:
\begin{itemize}
    \item proposals might be inserted by voters after the voting phase started.
    \item Voting might be extended as a permanent process. In other words, voting on and adding new proposals might be open continuously.
    \item A proposal is presented to a voter one by one as random outcomes of all proposals.
    \item Each vote might be of one and only type among: ``approve'', ``disapprove'' and ``indifferent''.
    \item Voters vote without authentication.
    \item Intended mostly for national rankings, but can also be local or have foreign participation.
    \item Should result in a ranking of proposals to assist public management.
\end{itemize}

This setting requires care about security and validity.
Some of which are:
\begin{itemize}
	\item adequate sampling of individual proposals and overall ranking.
    \item Registration of the IP address and time of votes to ease detection of automated and other fraudulent efforts.
    \item Reasonable use of the outcomes from the voting process. This requires probing the survey being conducted and its purposes. The indexes here presented target indicatives for the Brazilian federal government about the most important health care proposals. Given the unauthenticated voting, the outcomes might be regarded as reference rankings if data is minimally shared and checked for inadequate data entry (such as voting by automated scripts or a persistent participant introduced bias).
\end{itemize}

@@

The approval index $\alpha_i$ and the participation index $\gamma_i$ of the proposal $i$ was defined as:
\begin{equation}
\begin{split}
\alpha_i=\frac{v^+_i-v^-_i}{\eta_i} \\
\gamma_i=\frac{v^+_i+v^-_i}{\eta_i}
\end{split}
\end{equation}
where $v^+_i$, $v^-_i$ and $\eta_i$ are approval count, disapproval count and exhibition count, respectively.
Note that $\alpha_i\in[-1,1]$, $\gamma_i\in[0,1]$, and $v^o_i=\eta_i-v^+_i-v^-_i$ is the count of the ``indifferent'' manifestations received by proposal $i$.
Also, such $\alpha_i$ and $\gamma_i$ indexes are expected, for each proposal $i$, to be a constant  plus a sampling estimate error that should be smaller as $\eta_i$ raises. This error is thought to be acceptable if $\eta_i$ is above a threshold $\overline{\eta}$ established by the participation community and public managers.
As an initial decision, the staff agreed to use $\overline{\eta}$ as to select $10-20\%$ of all proposals.
A threshold $\overline{\gamma}$ can be used as a required level of engagement for proposals to be relevant,
while the threshold $\overline{\alpha}$ is used to classify the outcome
as ``approved'',``disapproved'' and ``clash''. More specifically:

@@

If a proposal is both sampled and relevant, than it is prioritized.
The coherent values of $\overline{\alpha_i}=0.5$ (or $1/3$)
and $\overline{\gamma_i}=0.5$ were chosen as standards of the decision model. 
These are likely to change with implementation and management.

@@

\vfill
\vspace{.6cm}
\vfill

@@

A Brazilian solution, used in diverse software and specially important as the output of a nation-wide decision making need, is the Agora Algorithm~\cite{agora}. It presents a decision procedure in phases (agenda proposition, deliberations proposition and commenting, voting) with resolution outcomes.
Although coherent, this framework requires authentication and might need experimentation and tuning in order to be effective with more than dozens or a few hundreds of participants.
